---
title: "Cell Annotation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cell Annotation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", fig.align = "center", out.width = "90%")
```

## Overview

This module finds cluster markers, visualizes protein panels, and assigns readable cell-type labels. The worked example starts from clustered **reg055_A** data and maps Seurat clusters to biological names.

## Load packages and object

```{r load, eval=FALSE}
library(Sphinx)
library(Seurat)
library(ggplot2)

obj <- readRDS("reg055_A_processed.rds")
```

## Top markers

```{r markers, eval=FALSE}
top5 <- find_top_markers(
  obj,
  assay = DefaultAssay(obj),
  save_path = "top5proteins.csv"
)
```

## Marker violin / heatmap / UMAP

```{r marker-viz, eval=FALSE}
plot_marker_violin(
  obj,
  markers = c("CD3", "CD20", "CD8", "CD4", "CD68", "Cytokeratin",
              "CD31", "aSMA", "CD45", "FOXP3", "Vimentin", "Ki67"),
  group_by = "seurat_clusters",
  assay = DefaultAssay(obj),
  save_path = "allmarker.png",
  width = 12, height = 14
)

plot_marker_heatmap(
  obj, top_markers = top5, group_by = "seurat_clusters",
  assay = DefaultAssay(obj), save_path = "marker_heatmap.png"
)

plot_umap_markers(
  obj, markers = head(top5$gene, 6),
  output_file = "umap_markers_plot.png"
)
```

```{r show-violin, echo=FALSE, out.width="90%", results='asis'}
f <- "allmarker.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r show-heat, echo=FALSE, out.width="85%", results='asis'}
f <- "marker_heatmap.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r show-umap-markers, echo=FALSE, out.width="85%", results='asis'}
f <- "umap_markers_plot.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r show-spatial-markers, echo=FALSE, out.width="85%", results='asis'}
f <- "spatial_markers.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Annotate and plot

Assign labels with `annotate_celltypes()` using a cluster-to-name map informed by top markers and the original CODEX `ClusterName` labels:

```{r annotate, eval=FALSE}
obj <- annotate_celltypes(
  obj,
  cluster_ids = c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"),
  celltype_labels = c(
    "CD4+ T cells", "CD8+ T cells", "CD68+CD163+ macrophages", "CD68+ macrophages",
    "tumor cells", "B cells", "tumor cells", "CD68+ macrophages", "B cells",
    "CD4+ T cells", "B cells", "B cells", "vasculature"
  ),
  cluster_column = "seurat_clusters"
)

plot_annotated_umap(obj, save_path = "umap_by_celltype.png")
plot_spatial_distribution(obj, save_path = "celltype_spatial_plot.png", point.size = 0.4)
```

```{r show-ann-umap, echo=FALSE, out.width="85%", results='asis'}
f <- "umap_by_celltype.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r show-spatial, echo=FALSE, out.width="85%", results='asis'}
f1 <- "celltype_spatial_plot.png"
f2 <- "01_spatial_distribution_celltype.png"
if (file.exists(f1)) knitr::include_graphics(f1) else if (file.exists(f2))
  knitr::include_graphics(f2) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

Colors stay consistent across UMAP, spatial, and network plots via `assign_celltype_colors()`.

## Save

```{r save, eval=FALSE}
saveRDS(obj, "reg055_A_annotated.rds")
```

## Next step

`vignette("spatial-network", package = "Sphinx")`
