---
title: "Cell Annotation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cell Annotation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This tutorial demonstrates how to identify cell types and annotate clusters in spatial proteomics data using Sphinx.

## Load Required Packages

```{r load-packages, eval=FALSE}
library(Sphinx)
library(Seurat)
library(ggplot2)
library(pheatmap)
library(dplyr)
```

## 1. Load Processed Data

```{r load-data, eval=FALSE}
# Load preprocessed spatial data
codex.obj <- readRDS("cycif1_processed.rds")

# Check object structure
print(codex.obj)
```

## 2. Identify Marker Proteins

```{r find-markers, eval=FALSE}
# Identify top marker proteins for each cluster
top5 <- find_top_markers(
  codex.obj, 
  assay = "RNA",
  save_path = "top5proteins.csv"
)

# View top markers
head(top5)
```

## 3. Visualize Marker Expression

### Violin Plots

```{r violin-plots, eval=FALSE}
# Create violin plots for all detected proteins
plot_marker_violin(
  codex.obj, 
  markers = rownames(codex.obj[["RNA"]]),
  assay = "RNA",
  save_path = "allmarker.png"
)
```


*Violin plots showing expression distribution of marker proteins across clusters.*

### Marker Heatmap

```{r marker-heatmap, eval=FALSE}
# Create heatmap of top marker expression
plot_marker_heatmap(
  codex.obj,
  top_markers = top5,
  save_path = "marker_heatmap.png"
)
```


*Heatmap plots showing expression of marker proteins across clusters.*

### Visualize marker expression in UMAP space

```{r umap-markers, eval=FALSE}
plot_umap_markers(codex.obj, 
                  markers = c("CD31", "CD19", "Pan-Cytokeratin",
                              "a-SMA","CD8","CD4"))
```


*Heatmap plots showing expression of marker proteins across clusters.*

### Visualize marker expression in spatial coordinates

```{r spatial-markers, eval=FALSE}
plot_spatial_markers(codex.obj, 
                  markers = c("CD31", "CD19", "Pan-Cytokeratin",
                              "a-SMA","CD8","CD4"),
                  point_size = 0.5)
```


*Heatmap plots showing spatial distribution of marker proteins across clusters.*

## 4. Cell Type Annotation

```{r annotate_celltypes, eval=FALSE}
# Define cluster to cell type mapping
cluster_ids <- c(0, 1, 2, 3, 4, 5, 6, 7锛?锛?                 9锛?0锛?1锛?2锛?3锛?4锛?5)
celltype_labels <- c(
  "Endothelial", "Tumor", "DC", "M2_Macrophage","Tumor","CAF","PD-1_CD4_T","B","Endothelial",
  "GC_B", "DC", "EMT_like_Tumor", "Proliferating_DC","Endothelial","Endothelial","Proliferating_Tumor"
)

# Apply cell type annotations
codex.obj <- annotate_celltypes(
  codex.obj,
  cluster_ids = cluster_ids,
  celltype_labels = celltype_labels,
  cluster_column = "RNA_snn_res.0.5"
)

# Check annotation results
table(codex.obj$celltype)
```

## 5. Visualize Annotated Results

### Annotated UMAP

```{r annotated-umap, eval=FALSE}
# Create UMAP colored by cell types
plot_annotated_umap(
  codex.obj,
  save_path = "umap_by_celltype.png"
)
```


*UMAP visualization colored by annotated cell types.*

### Spatial Distribution

```{r spatial-distribution, eval=FALSE}
# Visualize spatial distribution of cell types
plot_spatial_distribution(
  codex.obj,
  save_path = "celltype_spatial_plot.png"
)
```


*Spatial distribution of annotated cell types across tissue coordinates.*

## 6. Save Results

```{r save-results, eval=FALSE}
# Save annotated object
saveRDS(codex.obj, file = "tsu33_celltype.rds")

# Export metadata for downstream analysis
write.csv(codex.obj@meta.data, file = "tsu33_metadata.csv")
```
