---
title: "Functional Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functional Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", fig.align = "center", out.width = "92%")
```

## Overview

This module connects neighborhood structure to protein programs: differential testing (spatial-block aware by default), volcano plots with a capped y-axis, and enrichment summaries.

**Demo:** reg055_A panel expression with neighborhood clusters from the spatial-network tutorial.

## Prepare protein table

```{r prep, eval=FALSE}
library(Sphinx)
library(dplyr)

protein_df <- prepare_protein_data(clus, expr_df)
```

## Differential expression

```{r de, eval=FALSE}
diff_res <- perform_differential_expression(
  protein_df,
  test_level = "spatial_block"
)
```

## Volcano plots

`-log10(adjusted P)` is capped at `y_cap` (default **50**). Points above the cap are drawn as triangles on the truncation line.

```{r volcano, eval=FALSE}
plot_volcano_all_clusters(
  diff_res,
  diff_thresh = 0.15,
  p_thresh = 0.05,
  y_cap = 50,
  ncol = 3,
  save_plot = TRUE,
  output_dir = "functional_plots",
  filename = "VolcanoPlots_AllClusters"
)
```

```{r volcano-fig, echo=FALSE, results='asis'}
f <- "VolcanoPlots_AllClusters.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Enrichment

```{r enrich, eval=FALSE}
enrich <- perform_cluster_enrichment(
  diff_res,
  species = "human",
  pvalueCutoff = 0.05,
  mean_diff_cutoff = 0
)

plots <- plot_enrichment_results(
  enrich,
  top_n = 5,
  fdr_cutoff = 0.05,
  base_font_size = 12,
  save_plot = TRUE,
  output_dir = "enrichment_plots"
)
```

```{r bar-fig, echo=FALSE, results='asis'}
f <- "bar_plot.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r bubble-fig, echo=FALSE, results='asis'}
f <- "bubble_plot.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Tips for targeted CODEX panels

- Prefer the assayed protein list as EnrichR `background` (Sphinx default).
- Use `test_level = "spatial_block"` for tissue DE.
- Plotting helpers still work offline on any table with `Cluster`, `Term`, `FDR`, and `GenesN`.
