---
title: "Functional Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functional Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This tutorial demonstrates how to perform functional enrichment analysis on spatially defined neighborhoods to identify biological pathways and processes.

## Load Required Packages

```{r load-packages, eval=FALSE}
library(Sphinx)
library(ComplexHeatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(ggplot2)
library(ggrepel)
library(dplyr)
library(enrichR)
library(patchwork)
```

## 1. Load Data

```{r load-data, eval=FALSE}
# Load neighborhood clustering results
df <- read.csv("results.csv")

# Load protein expression data
protein <- read.csv("../tsu33_expression_filtered.csv", 
                   check.names = FALSE, 
                   row.names = 1)

# Check data dimensions
message("Neighborhood data: ", nrow(df), " cells")
message("Protein data: ", nrow(protein), " proteins x ", ncol(protein), " cells")
```

## 2. Prepare Protein Data

```{r prepare-data, eval=FALSE}
# Merge neighborhood clusters with protein expression
protein_df <- prepare_protein_data(df, protein)

# Check merged data structure
message("Merged data dimensions: ", paste(dim(protein_df), collapse = " x "))
message("NA values in key proteins: ", sum(is.na(protein_df$PD.1)))
```

## 3. Differential Expression Analysis

```{r diff-expression, eval=FALSE}
# Perform differential expression between neighborhoods
diff_results <- perform_differential_expression(protein_df)

# Save results
write.csv(diff_results, "diff_results.csv")

# View summary statistics
summary(diff_results$Log2FC)
table(diff_results$Significance)
```

## 4. Volcano Plot Visualization

```{r volcano-plots, eval=FALSE}
# Generate volcano plots for all clusters
volcano_all <- plot_volcano_all_clusters(diff_results)

# Adjust aspect ratio and save
volcano_all <- volcano_all + theme(aspect.ratio = 0.75)
ggsave("VolcanoPlots_AllClusters.pdf", volcano_all, width = 10, height = 20)
```


## 5. Functional Enrichment Analysis

```{r enrichment-analysis, eval=FALSE}
# Perform enrichment analysis using human databases
cluster_enrich <- perform_cluster_enrichment(
  diff_results, 
  species = "human"
)

# Save enrichment results
saveRDS(cluster_enrich, "cluster_enrich.rds")

# View enrichment summary
message("Enriched terms found: ", nrow(cluster_enrich))
message("Clusters with enrichment: ", length(unique(cluster_enrich$Cluster)))
```

**Enrichment Databases:** 
- GO Biological Processes 
- GO Molecular Functions
- KEGG Pathways 
- Reactome Pathways 
- CORUM Complexes

## 6. Visualize Enrichment Results

```{r visualize-enrichment, eval=FALSE}
# Generate publication-quality enrichment plots
plots <- plot_enrichment_results(
  cluster_enrich,
  top_n = 5,                   # Top 5 terms per cluster
  fdr_cutoff = 0.05,           # FDR threshold
  term_trunc_length = 40,      # Term name length
  base_font_size = 7,          # Font size
  plot_types = "bar",          # Plot type (bar, heatmap, dot)
  save_plot = TRUE             # Save plots automatically
)

# Access individual plots
bar_plot <- plots$bar_plot
heatmap_plot <- plots$heatmap
dot_plot <- plots$dot_plot
```

### Enrichment Bar Plot


### Enrichment bubble plot

