---
title: "Installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

## Requirements

- R >= 4.3
- For enrichment: `enrichR` (suggested); optionally `clusterProfiler` / `org.Hs.eg.db`
- Composition heatmaps use Bioconductor `ComplexHeatmap` / `circlize` (Imports)

## Install from GitHub

```{r github, eval=FALSE}
install.packages("devtools")
devtools::install_github("mongi126/Sphinx")
```

## Install from source tarball

```{r tarball, eval=FALSE}
install.packages("Sphinx_1.0.1.tar.gz", repos = NULL, type = "source")
```

## Load and check

```{r check, eval=FALSE}
library(Sphinx)
packageVersion("Sphinx")
help(package = "Sphinx")
```

## Example data in the docs

Tutorials use **reg055_A** from Schurch *et al.* CODEX imaging (`schurch2020coordinated`):

- Technology: CODEX
- ~3,887 cells, 58 protein channels
- Coordinates in `obsm/spatial`
- Published labels in `obs/ClusterName`

Any table with `Cell_ID`, `X`, `Y`, and a categorical annotation column works the same way.
