---
title: "Data Preprocessing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Preprocessing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", fig.align = "center", out.width = "90%")
```

## Overview

This module prepares spatial proteomics inputs (CODEX / CyCIF / MIBI tables or Seurat objects): QC filtering, normalization, dimensionality reduction, and clustering.

Examples use **reg055_A** (CODEX CRC, Schurch *et al.* 2020).

## Load packages

```{r pkgs, eval=FALSE}
library(Sphinx)
library(Seurat)
library(dplyr)
library(ggplot2)
```

## Load spatial data

```{r load, eval=FALSE}
obj <- load_spatial_data(filename = "reg055_A_counts_with_coords.csv")
# Or construct a Seurat object from an expression matrix plus X / Y metadata.
```

## QC filtering

```{r filter, eval=FALSE}
obj_filtered <- filter_data(obj, nFeature_quantile_threshold = 0.05)
```

## Normalize, PCA, cluster

```{r process, eval=FALSE}
obj_processed <- process_data(obj_filtered, dims = 1:15, resolution = 0.5)
```

## Elbow plot

```{r elbow, eval=FALSE}
plot_elbow(obj_processed, save_path = "elbow_plot.png")
```

```{r elbow-fig, echo=FALSE, out.width="70%", results='asis'}
f <- "elbow_plot.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Extract spatial coordinates

```{r coords, eval=FALSE}
obj_processed <- extract_spatial_coordinates(obj_processed)
```

## Quick cluster views

```{r viz, eval=FALSE}
plots <- visualize_results(obj_processed, save_dir = "preprocess_plots/")
# plots$umap_plot, plots$spatial_plot
```

```{r umap-fig, echo=FALSE, out.width="80%", results='asis'}
f <- "cluster_umap.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r spat-clus-fig, echo=FALSE, out.width="80%", results='asis'}
f <- "spatial_clusters.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

`visualize_results()` writes a cluster UMAP and a spatial map of cluster IDs using the package color palette.

## Save

```{r save, eval=FALSE}
saveRDS(obj_processed, "reg055_A_processed.rds")
```

## Next step

`vignette("annotation", package = "Sphinx")`, or jump to spatial networks if labels are ready: `vignette("spatial-network", package = "Sphinx")`.
