---
title: "Spatial Network Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial Network Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", fig.align = "center", out.width = "92%")
```

## Overview

This module builds adaptive spatial graphs, derives neighborhood features and clusters, and visualizes distances, composition, purity, and cell-cell interactions.

**Demo:** reg055_A (CODEX CRC, Schurch *et al.* 2020) with annotated cell-type labels from the annotation tutorial.

## Load packages and prepare data

```{r prep, eval=FALSE}
library(Sphinx)
library(data.table)
library(dplyr)
library(ggplot2)

# meta: Cell_ID, X, Y, celltype
df <- prepare_data(
  meta, cell_id_col = "Cell_ID", x_col = "X", y_col = "Y",
  celltype_col = "celltype"
)
```

## Spatial distribution

```{r spat, eval=FALSE}
visualize_spatial_distribution(
  df, point_size = 0.45, point_alpha = 0.95,
  title = "Annotated cell types"
)
```

```{r spat-fig, echo=FALSE, results='asis'}
f <- "01_spatial_distribution_celltype.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Cell-type distances

```{r dist, eval=FALSE}
dist_result <- calculate_celltype_distances(df, celltype_col = "celltype")
visualize_distance_heatmap(dist_result, show_values = TRUE)
visualize_distance_parallel(dist_result)
```

```{r dist-fig, echo=FALSE, results='asis'}
f <- "02_distance_heatmap.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Build spatial network

`build_spatial_network()` selects a graph method (`auto` / knn / radius / delaunay / window) and applies biological edge filtering (mutual nearest neighbors and distance gating where configured).

```{r net, eval=FALSE}
edges <- build_spatial_network(df, method = "auto", celltype_col = "celltype")
feat <- calculate_neighborhood_features(df, edges, celltype_col = "celltype")
clus <- cluster_neighborhoods(feat, edges, method = "kmeans", k = 10)
```

## Neighborhood clusters & composition

```{r cn, eval=FALSE}
visualize_spatial_distribution(
  clus, celltype_col = "Neighborhood_Cluster",
  point_size = 0.45, title = "Neighborhood clusters"
)

comp <- calculate_cluster_composition(clus)
plot_composition_barplot(comp)
plot_composition_heatmap(comp, cell_fontsize = 10)
```

```{r cn-fig, echo=FALSE, results='asis'}
f <- "05_spatial_distribution_clusters.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r bar-fig, echo=FALSE, results='asis'}
f <- "07_cluster_composition_barplot.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r heat-comp-fig, echo=FALSE, results='asis'}
f <- "06_composition_heatmap.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Neighborhood purity

```{r purity, eval=FALSE}
pur <- calculate_neighborhood_purity(clus, method = "knn", n_neighbors = 10)
visualize_neighborhood_purity(pur, point_size = 0.55)
```

```{r purity-fig, echo=FALSE, results='asis'}
f <- "04_neighborhood_purity.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Interactions

```{r intx, eval=FALSE}
intx <- analyze_spatial_interactions(clus, edges, celltype_col = "celltype")

visualize_interaction_heatmap(
  intx$interaction_matrix, transform = TRUE,
  display_numbers = FALSE, cellwidth = 10, cellheight = 10
)

visualize_interaction_network(intx$network)
```

```{r intx-hm, echo=FALSE, results='asis'}
f <- "08_interaction_heatmap.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r intx-net, echo=FALSE, results='asis'}
f <- "11_cell_interaction_network.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Spatial network graph (real edges)

Points are opaque with a white halo so colors remain clear over edges. For overview plots use `edge_mode = "top"`; for local zoom, shrink `zoom_radius` and downsample edges with `edge_mode = "random"` so structure stays readable.

```{r spatnet, eval=FALSE}
visualize_spatial_network(
  clus, edges, edge_mode = "top", top_n = 1200,
  point_size = 0.5, point_alpha = 1
)

cx <- mean(range(clus$X)); cy <- mean(range(clus$Y))
zr <- max(diff(range(clus$X)), diff(range(clus$Y))) * 0.07
visualize_spatial_network(
  clus, edges,
  edge_mode = "random", max_edges = 180,
  zoom_center = c(cx, cy), zoom_radius = zr,
  point_size = 2.2, point_alpha = 1,
  title = "Spatial network (local zoom)"
)
```

```{r spatnet-fig, echo=FALSE, results='asis'}
f <- "10_spatial_network.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

```{r spatnet-zoom, echo=FALSE, results='asis'}
f <- "10_spatial_network_zoom.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Voronoi view

```{r vor, eval=FALSE}
visualize_voronoi(clus, coloring = "neighborhood")
```

```{r vor-fig, echo=FALSE, results='asis'}
f <- "12_voronoi_neighborhood.png"
if (file.exists(f)) knitr::include_graphics(f) else
  cat("*Figure available on https://mongi126.github.io/Sphinx/*\n")
```

## Next step

`vignette("functional", package = "Sphinx")`
